Cost-effectiveness of a malaria elimination campaign

### 2017-02-01 paper.Rmd
finished preliminary effectiveness calculations. toma epidemiologos!

---
title: "Cost-effectiveness of a malaria elimination campaign"
author: "Laia Cirera, Joe Brew, Elisa Sicuri"
date: "1 de fevereiro de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
library(readxl)
library(tidyverse)

@@

# Read in incidence data
bes <- read_csv("data/cleaned/cases.csv")
```

Short formula in R: `r round(mean(bes$pk, na.rm=TRUE),digits=2)`

@@

Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investigação em Saúde de Manhiça (CISM). The main objective of this program is to assist the NMCP in adopting targeted evidence-based elimination plans through a 'learn by doing' strategy that begins in program mode and incorporates data from rigorous evaluation-into action as it is continuously being generated. Besides developing a fully costed, targeted, evidence-based elimination plan for Southern Mozambique, the first two years of this program have generated knowledge to inform the programmatic aspects of the elimination plan, most importantly how to interrupt transmission. It has also provided coordination, management and capacity building support to the NMCP, and facilitated the adoption of evidence-based malaria policies through the creation of a Malaria Technical Advisory Committee in the country. 
The  MALTEM program will provide technical, operational and financial support to design and scale up new strategies for the Mozambican National Malaria Control Program (NMCP) to achieve elimination in Maputo and significantly decrease malaria transmission rates in Gaza and Inhambane provinces by 2020.

@@

The malaria elimination is taking place in Magude District, on the North-Western area of Maputo Province, Southern Mozambique. Magude District borders with the district of Massingir (Gaza Province), in the North, Manhiça District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West. It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds. The district has seven health facilities with a total of 43 beds (1.65 beds per 1,000 inhabitants) and reported 20,000 malaria cases per 100,000 population at risk in 2012.

>>>>

---
title: "Cost-effectiveness of a malaria elimination campaign"
author: "Laia Cirera, Joe Brew, Elisa Sicuri"
date: "`r Sys.Date()`"
output:
 tufte::tufte_html: default
 tufte::tufte_handout:
   citation_package: natbib
   latex_engine: xelatex
 tufte::tufte_book:
   citation_package: natbib
   latex_engine: xelatex
bibliography: bibliography.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE)
```

```{r}
library(readxl)
library(tidyverse)
library(broom)

@@

# Read in incidence data and weatehr data
source("master.R")
bes <- df

# Create variables 
bes <- 
  bes %>%
  mutate(mda=ifelse(date<"2015-11-08", "pre_mda", 
                    ifelse(date<"2016-02-04", "mda", "post_mda")),
         intervention=district=="MAGUDE")

# Create model to estimate effectiveness
mod <-lm(p ~ mda*intervention + irs_coverage + age, data=bes, weights = population)
```

Model parameters

```{r}

# Explore results of linear model
# knitr::kable(tidy(summary(mod)))
```

```{r}

# Use model to predict conterfactual
other_irs <- bes %>% filter(year == 2016, district != "MAGUDE") %>% summarise(x =mean(irs_coverage, na.rm =TRUE)) %>% .$x
counter_factual <- 
   bes %>% filter(year == 2016, 
                  district == "MAGUDE")
counter_factual <- 
  counter_factual %>%
  mutate(irs_coverage = other_irs) %>%
  mutate(intervention = FALSE)

counter_factual$p_cf <- predict(mod, counter_factual)
counter_factual$cases_cf <- counter_factual$p_cf*counter_factual$population 
counter_factual$cases_averted <- counter_factual$cases_cf - counter_factual$cases

# Get sum of averted cases
ac <- counter_factual %>%
  group_by(age)%>%
  summarise(sum_cases_averted=sum(cases_averted))

# Get the reduction in incidence
inc <-
  counter_factual %>%
  group_by(age)%>%
  summarise(reduction = (mean(p_cf) - mean(p)) / mean(p_cf) * 100)

# Plot the observed and counterfactual
plot_data <- counter_factual %>%
  group_by(date) %>%
  summarise(cases = sum(cases),
            cases_cf = sum(cases_cf),
            cases_averted = sum(cases_averted)) %>%
  gather(key, value, cases:cases_averted) %>%
  filter(key != "cases_averted") %>%
  mutate(key = ifelse(key == "cases", "MALTEM", "Counter-factual"))

# library(ggthemes)
# ggplot(data=plot_data,
#        aes(x=date,
#            y=value,
#            color=key))+
#   geom_line(size =2,
#             alpha = 0.6) + 
#   labs(title="Observed malaria cases vs counterfactual",
#        subtitle="Post-MDA period, 2016",
#        x="Date",
#        y="Malaria cases") +
#   theme_minimal() +
#   scale_color_manual(name = "",
#                      values = c("red", "darkgreen"))

### Create new variables in our epi data related to clinical events

```

@@

Within this context and in line with the global and regional efforts towards malaria elimination, a five year malaria elimination program that started in 2014 has been established in Magude District (Mozambique) through Centro de Investigação em Saúde de Manhiça (CISM). The main objective of this program is to assist the NMCP in adopting targeted evidence-based elimination plans through a 'learn by doing' strategy that begins in program mode and incorporates data from rigorous evaluation-into action as it is continuously being generated. Besides developing a fully costed, targeted, evidence-based elimination plan for Southern Mozambique, the first two years of this program have generated knowledge to inform the programmatic aspects of the elimination plan, most importantly how to interrupt transmission. It has also provided coordination, management and capacity building support to the NMCP, and facilitated the adoption of evidence-based malaria policies through the creation of a Malaria Technical Advisory Committee in the country.

The  MALTEM program will provide technical, operational and financial support to design and scale up new strategies for the Mozambican National Malaria Control Program (NMCP) to achieve elimination in Maputo and significantly decrease malaria transmission rates in Gaza and Inhambane provinces by 2020.

@@

The malaria elimination is taking place in Magude District, on the North-Western area of Maputo Province, Southern Mozambique. Magude District borders with the district of Massingir (Gaza Province), in the North, Manhiça District and the province of Gaza (Chokwe and Bilene Districts) on the East, Moamba District in the South, and the South African Kruger National Park on the West. It has an area of 6,961 km2, with approximately 60,000 individuals and 11,408 family compounds. The district has seven health facilities with a total of 43 beds ^[(1.65 beds per 1,000 inhabitants)]  and reported 20,000 malaria cases per 100,000 population at risk in 2012.
